Detection of Treatment-Subgroup Interactions in Clustered Datasets with Generalized Linear Mixed-effects Model Trees (GLMM trees)

### 2015-08-31 pkg/glmertree/paper/treatment_interactions.Rnw
Re-added height/width arguments in DGP tree plot
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@766 255f4249-2263-4417-b1ea-8498779e27a7

\begin{figure}[!h]
\setkeys{Gin}{width=0.8\textwidth}
<<dgp-tree, fig=TRUE, height=5, width=7>>=
fig4 <- party(
partynode(1L,
 split = partysplit(2L, breaks = 30),
 kids = list(
   partynode(2L,
     split = partysplit(1L, breaks = 17),
     kids = list(
       partynode(3L, info = c(
  expression(''),
  expression(beta[j0] == '17.5'),
  expression(''),
  expression(beta[j1] == -'5.0'),
  expression(''),
  expression(d[j] == -'1.0')
 )),
       partynode(4L, info = c(
  expression(''),
  expression(beta[j0] == '30.0'),
  expression(''),
  expression(beta[j1] == '0.0'),
  expression(''),
  expression(d[j] == '0.0')
 )))),
   partynode(5L,
     split = partysplit(1L, breaks = 63),
     kids = list(
       partynode(6L, info = c(
  expression(''),
  expression(beta[j0] == '30.0'),
  expression(''),
  expression(beta[j1] == '0.0'),
  expression(''),
  expression(d[j] == '0.0')
 )),
       partynode(7L, info = c(
  expression(''),
  expression(beta[j0] == '42.5'),
  expression(''),
  expression(beta[j1] == '5.0'),
  expression(''),
  expression(d[j] == '1.0')
 )))))),
data.frame(U1 = numeric(0), U2 = numeric(0))
)
plot(fig4, tnex=1.5)
@
\caption{Data-generating model for treatment-subgroup interactions. Parameter $d$ denotes the standardized mean difference between the outcomes of Treatment 1 and 2 (i.e., $\beta_{j1} / \sigma_{\epsilon}$).}
\label{fig:modelC}
\end{figure}

@@

\begin{figure}[!htbp]
<<treesize-nointeract-xyplot, fig=TRUE, width=8, height=7>>=
load("treesizes_long_nointeract.dat")
treesizes.long$sigmabm <- factor(as.numeric(as.character(treesizes.long$sigmabm)), ordered=T)
treesizes.long$N <- factor(as.numeric(as.character(treesizes.long$N)), ordered=T)
aggdata <- aggregate(formula=treesize.values ~ treesize.ind + N + sigmabm + corUbm, FUN=mean, data=treesizes.long)
print(xyplot(treesize.values ~ sigmabm | N + corUbm, data = aggdata, groups=treesize.ind, type="b", ylab="tree size", xlab="sigma_b", par.settings=standard.theme("pdf",color=F), abline=c(1,0), auto.key=list(space="top", columns=2, title="Algorithm type", cex.title=1,lines=T, points=T)))
@
\caption{Average tree size of GLM and GLMM trees for datasets without treatment-subgroup interactions. Values $200$, $500$ and $1000$ refer to sample size. Reference line at $y=1$ represents the true tree size.}
\label{fig:xyplot_treesize_nointeract}
\end{figure}

@@

\begin{figure}[!htbp]
<<treesize-xyplot, fig=TRUE, width=9, height=9.5>>=
aggdata.size <- aggregate(formula=treesize.values ~ treesize.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)	
print(xyplot(treesize.values ~ sigmab | N + corUb, data = aggdata.size, groups=treesize.ind, type="b", ylab="tree size", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), abline=c(7,0), auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T)))
@
\caption{Average tree size of GLM and GLMM trees for datasets with treatment-subgroup interactions. Values $200$, $500$ and $1000$ refer to sample size. Reference line at $y=7$ represents true tree size.}
\label{fig:xyplot_treesize_interact}
\end{figure}

@@

\begin{figure}[!htbp]
<<accuracy-xyplot, fig=TRUE, width=9, height=9.5>>=
aggdata.acc <- aggregate(formula=truetree.values ~ truetree.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)
print(xyplot(truetree.values ~ sigmab | N + corUb, data = aggdata.acc, groups=truetree.ind, type="b",   ylab="tree accuracy", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T)))
@
\caption{Average accuracy of GLM and GLMM trees. Accuracy of trees is defined as the proportion of datasets in which the true tree was accurately recovered. Values are correlated to one of the $U_k$ variables; values $200$, $500$ and $1000$ refer to sample size.}
\label{fig:xyplot_treeaccuracy}
\end{figure}

@@

\begin{figure}[!htbp]
<<correlation-xyplot, fig=TRUE>>=
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + N + sigmab + treatdiff, FUN=mean, data=treespecs.long)
print(xyplot(correlation.values ~ sigmab | N + treatdiff, data = aggdata.cor, groups=correlation.ind, type="b", ylab="correlation", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T)))
@
\caption{Average predictive accuracy of GLM and GLMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted differences between Treatment 1 and 2. Values $5$ and $2.5$ refer to the absolute value of the unstandardized treatment-effect difference in subgroups with treatment-effect differences; values $200$, $500$ and $1000$ refer to sample size.}
\label{fig:xyplot_correlations}
\end{figure}

>>>>

\begin{figure}[!h]
\setkeys{Gin}{width=0.8\textwidth}
<<dgp-tree, fig=TRUE, height=5, width=7>>=
fig4 <- party(
  partynode(1L,
    split = partysplit(2L, breaks = 30),
    kids = list(
      partynode(2L,
        split = partysplit(1L, breaks = 17),
        kids = list(
          partynode(3L, info = c(
	    expression(''),
	    expression(beta[j0] == '17.5'),
	    expression(''),
	    expression(beta[j1] == -'5.0'),
	    expression(''),
	    expression(d[j] == -'1.0')
	  )),
          partynode(4L, info = c(
	    expression(''),
	    expression(beta[j0] == '30.0'),
	    expression(''),
	    expression(beta[j1] == '0.0'),
	    expression(''),
	    expression(d[j] == '0.0')
	  )))),
      partynode(5L,
        split = partysplit(1L, breaks = 63),
        kids = list(
          partynode(6L, info = c(
	    expression(''),
	    expression(beta[j0] == '30.0'),
	    expression(''),
	    expression(beta[j1] == '0.0'),
	    expression(''),
	    expression(d[j] == '0.0')
	  )),
          partynode(7L, info = c(
	    expression(''),
	    expression(beta[j0] == '42.5'),
	    expression(''),
	    expression(beta[j1] == '5.0'),
	    expression(''),
	    expression(d[j] == '1.0')
	  )))))),
  data.frame(U1 = numeric(0), U2 = numeric(0))
)
plot(fig4, tp_args = list(FUN = identity, width = 9), tnex = 1.5)
@
\caption{Data-generating model for treatment-subgroup interactions. Parameter $d$ denotes the standardized mean difference between the outcomes of Treatment 1 and 2 (i.e., $\beta_{j1} / \sigma_{\epsilon}$).}
\label{fig:modelC}
\end{figure}

@@

\begin{figure}[!htbp]
<<treesize-nointeract-xyplot, fig=TRUE, width=8, height=7>>=
load("treesizes_long_nointeract.dat")
treesizes.long$sigmabm <- factor(as.numeric(as.character(treesizes.long$sigmabm)), ordered=T)
treesizes.long$N <- factor(as.numeric(as.character(treesizes.long$N)), ordered=T)
aggdata <- aggregate(formula=treesize.values ~ treesize.ind + N + sigmabm + corUbm, FUN=mean, data=treesizes.long)
print(xyplot(treesize.values ~ sigmabm | N + corUbm, data = aggdata, groups=treesize.ind, 
	type="b", ylab="tree size", xlab="sigma_b", par.settings=standard.theme("pdf",color=F), 
	abline=c(1,0), auto.key=list(space="top", columns=2, title="Algorithm type", cex.title=1, 
	lines=T, points=T)))
@
\caption{Average tree size of GLM and GLMM trees for datasets without treatment-subgroup interactions. Values $200$, $500$ and $1000$ refer to sample size. Reference line at $y=1$ represents the true tree size.}
\label{fig:xyplot_treesize_nointeract}
\end{figure}

@@

\begin{figure}[!htbp]
<<treesize-xyplot, fig=TRUE, width=9, height=9.5>>=
aggdata.size <- aggregate(formula=treesize.values ~ treesize.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)	
print(xyplot(treesize.values ~ sigmab | N + corUb, data = aggdata.size, groups=treesize.ind, 
	type="b", ylab="tree size", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
	abline=c(7,0), auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
)
@
\caption{Average tree size of GLM and GLMM trees for datasets with treatment-subgroup interactions. Values $200$, $500$ and $1000$ refer to sample size. Reference line at $y=7$ represents true tree size.}
\label{fig:xyplot_treesize_interact}
\end{figure}

@@

\begin{figure}[!htbp]
<<accuracy-xyplot, fig=TRUE, width=9, height=9.5>>=
aggdata.acc <- aggregate(formula=truetree.values ~ truetree.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)
print(xyplot(truetree.values ~ sigmab | N + corUb, data = aggdata.acc, groups=truetree.ind, type="b", 
	ylab="tree accuracy", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
	auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
)
@
\caption{Average accuracy of GLM and GLMM trees. Accuracy of trees is defined as the proportion of datasets in which the true tree was accurately recovered. Values are correlated to one of the $U_k$ variables; values $200$, $500$ and $1000$ refer to sample size.}
\label{fig:xyplot_treeaccuracy}
\end{figure}

@@

\begin{figure}[!htbp]
<<correlation-xyplot, fig=TRUE>>=
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + N + sigmab + treatdiff, FUN=mean, data=treespecs.long)
print(xyplot(correlation.values ~ sigmab | N + treatdiff, data = aggdata.cor, groups=correlation.ind, 
	type="b", ylab="correlation", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
	auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
)
@
\caption{Average predictive accuracy of GLM and GLMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted differences between Treatment 1 and 2. Values $5$ and $2.5$ refer to the absolute value of the unstandardized treatment-effect difference in subgroups with treatment-effect differences; values $200$, $500$ and $1000$ refer to sample size.}
\label{fig:xyplot_correlations}
\end{figure}
